---
title: "Final Layer-Wise Hemisphere Analysis For SRM+LLM Encoding"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, message=FALSE, warning=FALSE}
# This Rmd organizes the layer-wise hemisphere analysis around three questions:
#
# Q1. After SRM reconstruction, which hemisphere shows stronger GPT-2 encoding?
# Q2. Which hemisphere shows larger SRM-related encoding gain?
# Q3. Does the LH/RH difference depend on GPT-2 layer depth?
#
# The code keeps the original logic: Welch independent tests are used first for
# LH-vs-RH parcel comparisons, and permutation tests are used as robustness checks.

options(scipen = 999)

required_packages <- c("tidyverse", "readr", "ggplot2", "writexl", "broom")
optional_packages <- c("lme4", "lmerTest", "broom.mixed")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    stop(paste("Required R package is not installed:", pkg))
  }
}
invisible(lapply(required_packages, library, character.only = TRUE))

HAS_LMER <- all(vapply(optional_packages, requireNamespace, quietly = TRUE, FUN.VALUE = logical(1)))
if (HAS_LMER) {
  invisible(lapply(optional_packages, library, character.only = TRUE))
}

BASE_DIR <- "YOUR_PROJECT_ROOT"

PARCEL_LAYER_PATH <- file.path(
  BASE_DIR,
  "parcel_layer_network_outputs",
  "step11A_build_parcel_level_encoding_table",
  "csv",
  "step11A_parcel_layer_encoding_table.csv"
)

PARCEL_GAIN_PATH <- file.path(
  BASE_DIR,
  "parcel_layer_network_outputs",
  "step11D_parcel_level_srm_gain_maps",
  "csv",
  "step11D_parcel_srm_gain_metrics.csv"
)

OUT_DIR <- file.path(BASE_DIR, "final_r_test_outputs")
TABLE_DIR <- file.path(OUT_DIR, "organized_tables")
FIGURE_DIR <- file.path(OUT_DIR, "organized_figures")

dir.create(TABLE_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(FIGURE_DIR, recursive = TRUE, showWarnings = FALSE)

NETWORK_ORDER <- c(
  "VisCent", "VisPeri",
  "SomMotA", "SomMotB",
  "DorsAttnA", "DorsAttnB",
  "SalVentAttnA", "SalVentAttnB",
  "LimbicA", "LimbicB",
  "ContA", "ContB", "ContC",
  "DefaultA", "DefaultB", "DefaultC",
  "TempPar"
)

LAYER_DEPTH_ORDER <- c("embedding", "early", "early_middle", "middle_late", "late")

MIN_N_PER_HEMI <- 4
N_PERM_LAYER_DEPTH <- 10000
N_PERM_EXACT_LAYER <- 10000
PERMUTATION_SEED <- 123
RUN_EXACT_LAYER_TESTS <- TRUE

fmt_num5 <- function(x) {
  ifelse(is.na(x), NA_character_, sprintf("%.5f", as.numeric(x)))
}

fmt_p5 <- function(p) {
  case_when(
    is.na(p) ~ NA_character_,
    p < 0.00001 ~ "<0.00001",
    TRUE ~ sprintf("%.5f", p)
  )
}

sig_stars <- function(p) {
  case_when(
    is.na(p) ~ "ns",
    p < 0.001 ~ "***",
    p < 0.01 ~ "**",
    p < 0.05 ~ "*",
    TRUE ~ "ns"
  )
}

add_note <- function(df, note) {
  df %>% mutate(analysis_note = note, .before = 1)
}

add_grouped_fdr <- function(df, group_cols, p_col = "p_value", out_col = "p_fdr") {
  df %>%
    group_by(across(all_of(group_cols))) %>%
    mutate("{out_col}" := ifelse(is.na(.data[[p_col]]), NA_real_, p.adjust(.data[[p_col]], method = "BH"))) %>%
    ungroup() %>%
    mutate(
      p_value_print = fmt_p5(.data[[p_col]]),
      p_fdr_print = fmt_p5(.data[[out_col]]),
      significance = sig_stars(.data[[out_col]]),
      fdr_family = paste(group_cols, collapse = " + ")
    )
}

cohens_d_independent <- function(left_vals, right_vals) {
  left_vals <- left_vals[is.finite(left_vals)]
  right_vals <- right_vals[is.finite(right_vals)]
  n_left <- length(left_vals)
  n_right <- length(right_vals)
  if (n_left < 2 || n_right < 2) return(NA_real_)
  pooled_sd <- sqrt(
    ((n_left - 1) * stats::var(left_vals) + (n_right - 1) * stats::var(right_vals)) /
      (n_left + n_right - 2)
  )
  if (!is.finite(pooled_sd) || pooled_sd == 0) return(NA_real_)
  (mean(left_vals) - mean(right_vals)) / pooled_sd
}

safe_welch_test <- function(df, value_col, group_col = "hemisphere") {
  dat <- df %>%
    transmute(
      value = as.numeric(.data[[value_col]]),
      group = as.factor(.data[[group_col]])
    ) %>%
    filter(is.finite(value), !is.na(group))

  left_vals <- dat %>% filter(group == "LH") %>% pull(value)
  right_vals <- dat %>% filter(group == "RH") %>% pull(value)

  base_row <- tibble(
    value_col = value_col,
    n_left = length(left_vals),
    n_right = length(right_vals),
    mean_left = ifelse(length(left_vals) > 0, mean(left_vals), NA_real_),
    mean_right = ifelse(length(right_vals) > 0, mean(right_vals), NA_real_),
    median_left = ifelse(length(left_vals) > 0, median(left_vals), NA_real_),
    median_right = ifelse(length(right_vals) > 0, median(right_vals), NA_real_),
    sd_left = ifelse(length(left_vals) > 1, sd(left_vals), NA_real_),
    sd_right = ifelse(length(right_vals) > 1, sd(right_vals), NA_real_),
    mean_difference_left_minus_right = ifelse(
      length(left_vals) > 0 && length(right_vals) > 0,
      mean(left_vals) - mean(right_vals),
      NA_real_
    ),
    cohen_d_left_minus_right = cohens_d_independent(left_vals, right_vals),
    test_method = "two-sided Welch independent t-test"
  )

  if (length(left_vals) < MIN_N_PER_HEMI || length(right_vals) < MIN_N_PER_HEMI) {
    return(base_row %>%
      mutate(
        statistic = NA_real_,
        df = NA_real_,
        p_value = NA_real_,
        note = paste0("Skipped: fewer than ", MIN_N_PER_HEMI, " parcels in at least one hemisphere")
      ))
  }

  fit <- stats::t.test(left_vals, right_vals, paired = FALSE, alternative = "two.sided")

  base_row %>%
    mutate(
      statistic = unname(fit$statistic),
      df = unname(fit$parameter),
      p_value = unname(fit$p.value),
      note = NA_character_
    )
}

safe_permutation_test <- function(df, value_col, group_col = "hemisphere", n_perm = N_PERM_LAYER_DEPTH) {
  dat <- df %>%
    transmute(
      value = as.numeric(.data[[value_col]]),
      group = as.factor(.data[[group_col]])
    ) %>%
    filter(is.finite(value), !is.na(group))

  left_vals <- dat %>% filter(group == "LH") %>% pull(value)
  right_vals <- dat %>% filter(group == "RH") %>% pull(value)

  base_row <- tibble(
    value_col = value_col,
    n_left = length(left_vals),
    n_right = length(right_vals),
    mean_left = ifelse(length(left_vals) > 0, mean(left_vals), NA_real_),
    mean_right = ifelse(length(right_vals) > 0, mean(right_vals), NA_real_),
    median_left = ifelse(length(left_vals) > 0, median(left_vals), NA_real_),
    median_right = ifelse(length(right_vals) > 0, median(right_vals), NA_real_),
    sd_left = ifelse(length(left_vals) > 1, sd(left_vals), NA_real_),
    sd_right = ifelse(length(right_vals) > 1, sd(right_vals), NA_real_),
    mean_difference_left_minus_right = ifelse(
      length(left_vals) > 0 && length(right_vals) > 0,
      mean(left_vals) - mean(right_vals),
      NA_real_
    ),
    cohen_d_left_minus_right = cohens_d_independent(left_vals, right_vals),
    test_method = "two-sided label-permutation mean-difference test",
    n_perm = n_perm
  )

  if (length(left_vals) < MIN_N_PER_HEMI || length(right_vals) < MIN_N_PER_HEMI) {
    return(base_row %>%
      mutate(
        p_value = NA_real_,
        note = paste0("Skipped: fewer than ", MIN_N_PER_HEMI, " parcels in at least one hemisphere")
      ))
  }

  observed_diff <- mean(left_vals) - mean(right_vals)
  all_vals <- c(left_vals, right_vals)
  n_left <- length(left_vals)

  perm_diffs <- replicate(n_perm, {
    shuffled <- sample(all_vals, length(all_vals), replace = FALSE)
    mean(shuffled[seq_len(n_left)]) - mean(shuffled[(n_left + 1):length(shuffled)])
  })

  p_perm <- (sum(abs(perm_diffs) >= abs(observed_diff)) + 1) / (n_perm + 1)

  base_row %>%
    mutate(
      p_value = p_perm,
      note = NA_character_
    )
}

format_report_sentence <- function(row) {
  if (is.na(row[["p_value"]])) return(NA_character_)
  paste0(
    row[["roi_name"]], " | ", row[["metric_label"]], " | ", row[["layer_depth"]],
    ": LH-RH = ", fmt_num5(row[["mean_difference_left_minus_right"]]),
    ", t(", fmt_num5(row[["df"]]), ") = ", fmt_num5(row[["statistic"]]),
    ", p = ", fmt_p5(row[["p_value"]]),
    ", FDR p = ", fmt_p5(row[["p_fdr"]]), "."
  )
}
```

# 1. Load And Validate Inputs

```{r load-validate}
if (!file.exists(PARCEL_LAYER_PATH)) stop(paste("Missing parcel-layer input:", PARCEL_LAYER_PATH))
if (!file.exists(PARCEL_GAIN_PATH)) stop(paste("Missing parcel-gain input:", PARCEL_GAIN_PATH))

parcel_layer_df <- readr::read_csv(PARCEL_LAYER_PATH, show_col_types = FALSE)
parcel_gain_df <- readr::read_csv(PARCEL_GAIN_PATH, show_col_types = FALSE)

required_layer_cols <- c(
  "roi_name", "network_family", "parcel_id", "parcel_label", "hemisphere",
  "layer_index", "layer_label", "log_layer_depth_percent", "layer_depth_bin",
  "raw_mean_r", "srm_mean_r", "delta_mean_r"
)
missing_layer_cols <- setdiff(required_layer_cols, names(parcel_layer_df))
if (length(missing_layer_cols) > 0) {
  stop(paste("Missing columns in parcel_layer_df:", paste(missing_layer_cols, collapse = ", ")))
}

required_gain_cols <- c(
  "roi_name", "network_family", "parcel_id", "parcel_label", "hemisphere",
  "best_srm_r", "same_layer_gain", "overall_mean_gain", "best_srm_log_depth_percent",
  "best_srm_depth_bin"
)
missing_gain_cols <- setdiff(required_gain_cols, names(parcel_gain_df))
if (length(missing_gain_cols) > 0) {
  stop(paste("Missing columns in parcel_gain_df:", paste(missing_gain_cols, collapse = ", ")))
}

clean_hemi <- function(x) {
  case_when(
    x %in% c("LH", "left", "Left", "L", "l") ~ "LH",
    x %in% c("RH", "right", "Right", "R", "r") ~ "RH",
    TRUE ~ as.character(x)
  )
}

available_layer_depths <- unique(as.character(parcel_layer_df$layer_depth_bin))
layer_depth_order_used <- LAYER_DEPTH_ORDER[LAYER_DEPTH_ORDER %in% available_layer_depths]
if (length(layer_depth_order_used) == 0) {
  stop("No recognized layer-depth categories were found in parcel_layer_df$layer_depth_bin.")
}

parcel_layer_df <- parcel_layer_df %>%
  mutate(
    roi_name = factor(roi_name, levels = NETWORK_ORDER),
    hemisphere = factor(clean_hemi(hemisphere), levels = c("LH", "RH")),
    layer_index = as.integer(layer_index),
    layer_depth = factor(layer_depth_bin, levels = layer_depth_order_used),
    log_layer_depth_percent = as.numeric(log_layer_depth_percent),
    parcel_id = as.factor(parcel_id),
    raw_mean_r = as.numeric(raw_mean_r),
    srm_mean_r = as.numeric(srm_mean_r),
    delta_mean_r = as.numeric(delta_mean_r)
  )

parcel_gain_df <- parcel_gain_df %>%
  mutate(
    roi_name = factor(roi_name, levels = NETWORK_ORDER),
    hemisphere = factor(clean_hemi(hemisphere), levels = c("LH", "RH")),
    best_srm_layer_depth = factor(best_srm_depth_bin, levels = layer_depth_order_used),
    parcel_id = as.factor(parcel_id),
    best_srm_r = as.numeric(best_srm_r),
    same_layer_gain = as.numeric(same_layer_gain),
    overall_mean_gain = as.numeric(overall_mean_gain),
    best_srm_log_depth_percent = as.numeric(best_srm_log_depth_percent)
  )

critical_cols <- c("raw_mean_r", "srm_mean_r", "delta_mean_r", "log_layer_depth_percent")
if (any(!is.finite(as.matrix(parcel_layer_df[critical_cols])))) {
  stop("Non-finite values detected in critical parcel-layer numeric columns.")
}

hemisphere_counts <- parcel_gain_df %>%
  distinct(roi_name, network_family, parcel_id, hemisphere) %>%
  count(roi_name, network_family, hemisphere, name = "n_parcels") %>%
  tidyr::pivot_wider(names_from = hemisphere, values_from = n_parcels, values_fill = 0) %>%
  mutate(
    n_total = LH + RH,
    min_n_per_hemi = pmin(LH, RH),
    sufficient_for_independent_tests = min_n_per_hemi >= MIN_N_PER_HEMI
  ) %>%
  arrange(factor(roi_name, levels = NETWORK_ORDER)) %>%
  add_note(
    "This table checks how many left- and right-hemisphere parcels are available in each network. It matters because all LH/RH parcel tests are independent tests and require enough parcels in both hemispheres."
  )

input_sanity_summary <- tibble(
  n_rows_parcel_layer = nrow(parcel_layer_df),
  n_unique_parcels = n_distinct(parcel_layer_df$parcel_id),
  n_networks = n_distinct(parcel_layer_df$roi_name),
  n_layers = n_distinct(parcel_layer_df$layer_index),
  n_layer_depth_categories = n_distinct(parcel_layer_df$layer_depth),
  layer_depth_categories = paste(layer_depth_order_used, collapse = ", "),
  n_rows_parcel_gain = nrow(parcel_gain_df),
  has_lmer = HAS_LMER,
  exact_layer_tests_enabled = RUN_EXACT_LAYER_TESTS
) %>%
  add_note(
    "This table verifies that the input data contain the expected parcel-layer rows, networks, GPT-2 layers, and layer-depth categories before running hemisphere statistics."
  )

input_sanity_summary
hemisphere_counts
```

# 2. Define Outcomes And Analysis Questions

```{r define-outcomes}
metric_labels <- c(
  delta_mean_r = "SRM-related encoding gain (SRM+LLM - Raw+LLM)",
  srm_mean_r = "SRM+LLM encoding r",
  raw_mean_r = "Raw+LLM encoding r"
)

question_table <- tribble(
  ~analysis_question, ~primary_metric, ~statistical_test, ~interpretation,
  "Q1: Which hemisphere shows stronger SRM+LLM encoding?", "srm_mean_r",
  "Welch/permutation tests within layer depth, plus mixed-effects models across continuous layer depth.",
  "This asks whether left- or right-hemisphere parcels are more predictable from GPT-2 after SRM reconstruction.",
  "Q2: Which hemisphere shows larger SRM-related encoding gain?", "delta_mean_r",
  "Welch/permutation tests within layer depth, plus mixed-effects models across continuous layer depth.",
  "This asks whether SRM improves LLM encoding more strongly in left- or right-hemisphere parcels.",
  "Q3: Does the hemisphere effect depend on GPT-2 layer depth?", "delta_mean_r and srm_mean_r",
  "Network-wise mixed-effects models and layer-depth follow-up tests.",
  "This asks whether LH/RH differences change from embedding/early layers to middle-late/late GPT-2 representations."
) %>%
  add_note(
    "This sheet is the roadmap for the Rmd. The final tables are ordered around these three questions rather than around implementation details."
  )

outcome_definition <- tibble(
  variable = names(metric_labels),
  label = unname(metric_labels),
  meaning = c(
    "The gain attributable to SRM reconstruction at a given parcel and GPT-2 layer.",
    "The final encoding strength when GPT-2 predicts SRM-reconstructed responses.",
    "The baseline encoding strength when GPT-2 predicts raw anatomically aligned responses."
  )
) %>%
  add_note(
    "This table defines the variables used in the statistical tests. The main outcome is delta_mean_r; srm_mean_r is the final predictability measure; raw_mean_r is retained as baseline context."
  )

question_table
outcome_definition
```

# 3. Q1 And Q2: LH/RH Profiles Across GPT-2 Layers

```{r profiles, fig.width=16, fig.height=10}
profile_df <- parcel_layer_df %>%
  group_by(roi_name, network_family, hemisphere, layer_index, layer_label, layer_depth) %>%
  summarise(
    raw_mean_r = mean(raw_mean_r, na.rm = TRUE),
    srm_mean_r = mean(srm_mean_r, na.rm = TRUE),
    delta_mean_r = mean(delta_mean_r, na.rm = TRUE),
    n_parcels = n_distinct(parcel_id),
    .groups = "drop"
  ) %>%
  mutate(roi_name = factor(roi_name, levels = NETWORK_ORDER)) %>%
  add_note(
    "This table averages parcel-level values within each network, hemisphere, and GPT-2 layer. It visualizes whether LH/RH differences are broad trends or restricted to specific layer depths."
  )

profile_long_df <- profile_df %>%
  select(-analysis_note) %>%
  pivot_longer(
    cols = all_of(c("delta_mean_r", "srm_mean_r")),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(metric_label = recode(metric, !!!metric_labels))

p_delta_profiles <- profile_long_df %>%
  filter(metric == "delta_mean_r") %>%
  ggplot(aes(x = layer_index, y = value, color = hemisphere)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60", linewidth = 0.35) +
  geom_line(linewidth = 0.8, alpha = 0.9) +
  facet_wrap(~ roi_name, ncol = 5, scales = "fixed") +
  scale_color_manual(values = c("LH" = "#D55E00", "RH" = "#0072B2")) +
  labs(
    title = "Q2: LH/RH SRM-Related Encoding Gain Across GPT-2 Layers",
    x = "GPT-2 representation index (0 = embedding; 1-48 = transformer layers)",
    y = "Mean gain in encoding r (SRM+LLM - Raw+LLM)",
    color = "Hemisphere"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right", strip.text = element_text(face = "bold"))

p_delta_profiles

ggsave(
  filename = file.path(FIGURE_DIR, "Q2_lh_rh_srm_gain_layer_profiles.png"),
  plot = p_delta_profiles,
  width = 16,
  height = 10,
  dpi = 300
)

p_srm_profiles <- profile_long_df %>%
  filter(metric == "srm_mean_r") %>%
  ggplot(aes(x = layer_index, y = value, color = hemisphere)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60", linewidth = 0.35) +
  geom_line(linewidth = 0.8, alpha = 0.9) +
  facet_wrap(~ roi_name, ncol = 5, scales = "fixed") +
  scale_color_manual(values = c("LH" = "#D55E00", "RH" = "#0072B2")) +
  labs(
    title = "Q1: LH/RH SRM+LLM Encoding Across GPT-2 Layers",
    x = "GPT-2 representation index (0 = embedding; 1-48 = transformer layers)",
    y = "Mean SRM+LLM encoding r",
    color = "Hemisphere"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right", strip.text = element_text(face = "bold"))

p_srm_profiles

ggsave(
  filename = file.path(FIGURE_DIR, "Q1_lh_rh_srm_encoding_layer_profiles.png"),
  plot = p_srm_profiles,
  width = 16,
  height = 10,
  dpi = 300
)
```

# 4. Q3: Network-Wise Mixed-Effects Models

```{r mixed-models}
if (!HAS_LMER) {
  layer_model_results <- tibble(
    analysis_note = "Mixed-effects models were skipped because lme4, lmerTest, or broom.mixed is not installed.",
    model_question = NA_character_,
    value_col = NA_character_,
    roi_name = NA_character_,
    network_family = NA_character_,
    effect = NA_character_,
    NumDF = NA_real_,
    DenDF = NA_real_,
    F_value = NA_real_,
    p_value = NA_real_,
    p_fdr = NA_real_,
    significance = NA_character_
  )
} else {
  run_network_lmer <- function(df, metric_name) {
    dat <- df %>%
      transmute(
        value = as.numeric(.data[[metric_name]]),
        hemisphere = factor(hemisphere, levels = c("RH", "LH")),
        log_layer_depth_percent = as.numeric(log_layer_depth_percent),
        parcel_id = as.factor(parcel_id)
      ) %>%
      filter(is.finite(value), is.finite(log_layer_depth_percent), !is.na(hemisphere))

    if (nrow(dat) < 20 || n_distinct(dat$hemisphere) < 2 || n_distinct(dat$parcel_id) < 4) {
      return(tibble(effect = NA_character_, NumDF = NA_real_, DenDF = NA_real_, F_value = NA_real_, p_value = NA_real_))
    }

    fit <- lmerTest::lmer(value ~ hemisphere * scale(log_layer_depth_percent) + (1 | parcel_id), data = dat)
    an <- as.data.frame(anova(fit, type = 3)) %>%
      rownames_to_column("effect") %>%
      filter(effect != "(Intercept)")

    an %>%
      transmute(
        effect = effect,
        NumDF = .data[["NumDF"]],
        DenDF = .data[["DenDF"]],
        F_value = .data[["F value"]],
        p_value = .data[["Pr(>F)"]]
      )
  }

  layer_model_results <- purrr::map_dfr(c("delta_mean_r", "srm_mean_r"), function(metric_name) {
    parcel_layer_df %>%
      group_by(roi_name, network_family) %>%
      group_modify(~ run_network_lmer(.x, metric_name)) %>%
      ungroup() %>%
      mutate(value_col = metric_name)
  }) %>%
    add_grouped_fdr(group_cols = c("value_col", "effect")) %>%
    mutate(
      metric_label = recode(value_col, !!!metric_labels),
      model_question = case_when(
        value_col == "srm_mean_r" ~ "Q1/Q3: final SRM+LLM encoding and its layer-depth dependence",
        value_col == "delta_mean_r" ~ "Q2/Q3: SRM-related gain and its layer-depth dependence",
        TRUE ~ "other"
      ),
      effect_label = recode(
        effect,
        hemisphere = "Hemisphere main effect",
        `scale(log_layer_depth_percent)` = "Layer depth main effect",
        `hemisphere:scale(log_layer_depth_percent)` = "Hemisphere x layer depth interaction",
        .default = effect
      )
    ) %>%
    arrange(value_col, effect, p_fdr, roi_name) %>%
    add_note(
      "This table is the main model-based answer to Q3. Each network is modeled separately with parcel as a random intercept; the hemisphere x layer-depth interaction tests whether LH/RH differences change across GPT-2 representational depth."
    )
}

layer_model_results
```

# 5. Q1 And Q2 Follow-Up: Layer-Depth Welch Tests

```{r layer-depth-welch}
parcel_layer_depth_df <- parcel_layer_df %>%
  group_by(roi_name, network_family, parcel_id, parcel_label, hemisphere, layer_depth) %>%
  summarise(
    raw_mean_r = mean(raw_mean_r, na.rm = TRUE),
    srm_mean_r = mean(srm_mean_r, na.rm = TRUE),
    delta_mean_r = mean(delta_mean_r, na.rm = TRUE),
    n_layers_in_depth = dplyr::n(),
    mean_log_layer_depth_percent = mean(log_layer_depth_percent, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    roi_name = factor(roi_name, levels = NETWORK_ORDER),
    layer_depth = factor(layer_depth, levels = layer_depth_order_used)
  )

layer_depth_welch_tests <- purrr::map_dfr(c("delta_mean_r", "srm_mean_r"), function(metric_name) {
  parcel_layer_depth_df %>%
    group_by(roi_name, network_family, layer_depth) %>%
    group_modify(~ safe_welch_test(.x, value_col = metric_name, group_col = "hemisphere")) %>%
    ungroup()
}) %>%
  add_grouped_fdr(group_cols = c("value_col", "layer_depth")) %>%
  mutate(
    metric_label = recode(value_col, !!!metric_labels),
    direction = case_when(
      is.na(mean_difference_left_minus_right) ~ "undefined",
      mean_difference_left_minus_right > 0 ~ "LH > RH",
      mean_difference_left_minus_right < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    ),
    report_sentence = pmap_chr(
      list(roi_name, metric_label, layer_depth, mean_difference_left_minus_right, df, statistic, p_value, p_fdr),
      function(roi_name, metric_label, layer_depth, diff, df, statistic, p_value, p_fdr) {
        if (is.na(p_value)) return(NA_character_)
        paste0(
          roi_name, " | ", metric_label, " | ", layer_depth,
          ": LH-RH = ", fmt_num5(diff),
          ", t(", fmt_num5(df), ") = ", fmt_num5(statistic),
          ", p = ", fmt_p5(p_value),
          ", FDR p = ", fmt_p5(p_fdr), "."
        )
      }
    )
  ) %>%
  arrange(value_col, layer_depth, p_fdr, roi_name) %>%
  add_note(
    "This table performs the first follow-up for Q1 and Q2: two-tailed Welch independent tests comparing LH parcels with RH parcels within each network and layer-depth category. Each parcel is averaged within a layer depth first, so repeated layers from the same parcel are not treated as independent."
  )

layer_depth_welch_tests
```

# 6. Q1 And Q2 Follow-Up: Layer-Depth Permutation Tests

```{r layer-depth-permutation}
set.seed(PERMUTATION_SEED)

layer_depth_permutation_tests <- purrr::map_dfr(c("delta_mean_r", "srm_mean_r"), function(metric_name) {
  parcel_layer_depth_df %>%
    group_by(roi_name, network_family, layer_depth) %>%
    group_modify(~ safe_permutation_test(.x, value_col = metric_name, group_col = "hemisphere", n_perm = N_PERM_LAYER_DEPTH)) %>%
    ungroup()
}) %>%
  add_grouped_fdr(group_cols = c("value_col", "layer_depth")) %>%
  mutate(
    metric_label = recode(value_col, !!!metric_labels),
    direction = case_when(
      is.na(mean_difference_left_minus_right) ~ "undefined",
      mean_difference_left_minus_right > 0 ~ "LH > RH",
      mean_difference_left_minus_right < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    )
  ) %>%
  arrange(value_col, layer_depth, p_fdr, roi_name) %>%
  add_note(
    "This table repeats the layer-depth LH/RH comparison with two-sided label permutations. It is used as the non-parametric robustness check after the Welch tests."
  )

layer_depth_permutation_tests
```

# 7. Exploratory Exact-Layer Localization

```{r exact-layer-tests}
set.seed(PERMUTATION_SEED)

if (RUN_EXACT_LAYER_TESTS) {
  exact_layer_welch_tests <- purrr::map_dfr(c("delta_mean_r", "srm_mean_r"), function(metric_name) {
    parcel_layer_df %>%
      group_by(roi_name, network_family, layer_index, layer_label, layer_depth) %>%
      group_modify(~ safe_welch_test(.x, value_col = metric_name, group_col = "hemisphere")) %>%
      ungroup()
  }) %>%
    add_grouped_fdr(group_cols = c("value_col", "layer_index")) %>%
    mutate(
      metric_label = recode(value_col, !!!metric_labels),
      direction = case_when(
        is.na(mean_difference_left_minus_right) ~ "undefined",
        mean_difference_left_minus_right > 0 ~ "LH > RH",
        mean_difference_left_minus_right < 0 ~ "RH > LH",
        TRUE ~ "LH = RH"
      )
    ) %>%
    arrange(value_col, layer_index, p_fdr, roi_name) %>%
    add_note(
      "This exploratory table uses Welch tests at each exact GPT-2 layer to localize which layers may drive the broader layer-depth effects. It is multiple-comparison-heavy and should be interpreted as localization rather than the main test."
    )

  exact_layer_permutation_tests <- purrr::map_dfr(c("delta_mean_r", "srm_mean_r"), function(metric_name) {
    parcel_layer_df %>%
      group_by(roi_name, network_family, layer_index, layer_label, layer_depth) %>%
      group_modify(~ safe_permutation_test(.x, value_col = metric_name, group_col = "hemisphere", n_perm = N_PERM_EXACT_LAYER)) %>%
      ungroup()
  }) %>%
    add_grouped_fdr(group_cols = c("value_col", "layer_index")) %>%
    mutate(
      metric_label = recode(value_col, !!!metric_labels),
      direction = case_when(
        is.na(mean_difference_left_minus_right) ~ "undefined",
        mean_difference_left_minus_right > 0 ~ "LH > RH",
        mean_difference_left_minus_right < 0 ~ "RH > LH",
        TRUE ~ "LH = RH"
      )
    ) %>%
    arrange(value_col, layer_index, p_fdr, roi_name) %>%
    add_note(
      "This exploratory table repeats exact-layer localization with two-sided permutation tests. It provides a robustness check for exact layers identified by Welch tests."
    )
} else {
  exact_layer_welch_tests <- tibble(
    analysis_note = "Exact-layer Welch tests were skipped. Set RUN_EXACT_LAYER_TESTS <- TRUE to run them."
  )
  exact_layer_permutation_tests <- tibble(
    analysis_note = "Exact-layer permutation tests were skipped. Set RUN_EXACT_LAYER_TESTS <- TRUE to run them."
  )
}

exact_layer_welch_tests %>% filter(!is.na(p_fdr), p_fdr < 0.05) %>% head(30)
exact_layer_permutation_tests %>% filter(!is.na(p_fdr), p_fdr < 0.05) %>% head(30)
```

# 8. Direction Summaries

```{r direction-summaries}
layer_depth_direction_summary <- parcel_layer_depth_df %>%
  group_by(roi_name, network_family, layer_depth, hemisphere) %>%
  summarise(
    mean_delta_r = mean(delta_mean_r, na.rm = TRUE),
    mean_srm_r = mean(srm_mean_r, na.rm = TRUE),
    n_parcels = n_distinct(parcel_id),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = hemisphere,
    values_from = c(mean_delta_r, mean_srm_r, n_parcels),
    names_sep = "_"
  ) %>%
  mutate(
    delta_lh_minus_rh = mean_delta_r_LH - mean_delta_r_RH,
    srm_lh_minus_rh = mean_srm_r_LH - mean_srm_r_RH,
    delta_direction = case_when(
      is.na(delta_lh_minus_rh) ~ "undefined",
      delta_lh_minus_rh > 0 ~ "LH > RH",
      delta_lh_minus_rh < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    ),
    srm_direction = case_when(
      is.na(srm_lh_minus_rh) ~ "undefined",
      srm_lh_minus_rh > 0 ~ "LH > RH",
      srm_lh_minus_rh < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    )
  ) %>%
  add_note(
    "This table summarizes the direction of LH/RH differences within each network and layer depth. Positive values mean LH > RH; negative values mean RH > LH."
  )

network_direction_summary <- layer_depth_direction_summary %>%
  select(-analysis_note) %>%
  group_by(roi_name, network_family) %>%
  summarise(
    mean_delta_lh_minus_rh_across_depths = mean(delta_lh_minus_rh, na.rm = TRUE),
    mean_srm_lh_minus_rh_across_depths = mean(srm_lh_minus_rh, na.rm = TRUE),
    n_delta_lh_biased_depths = sum(delta_lh_minus_rh > 0, na.rm = TRUE),
    n_delta_rh_biased_depths = sum(delta_lh_minus_rh < 0, na.rm = TRUE),
    n_depths = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_delta_lh_minus_rh_across_depths)) %>%
  add_note(
    "This table compresses the layer-depth direction summary to the network level. It is descriptive and helps interpret whether a network is consistently LH- or RH-biased across layer depth."
  )

layer_depth_direction_summary
network_direction_summary
```

# 9. Figures

```{r figures, fig.width=14, fig.height=7}
heatmap_df <- layer_depth_permutation_tests %>%
  filter(value_col %in% c("delta_mean_r", "srm_mean_r")) %>%
  mutate(
    metric_label = recode(value_col, !!!metric_labels),
    roi_name = factor(roi_name, levels = NETWORK_ORDER),
    layer_depth = factor(layer_depth, levels = layer_depth_order_used)
  )

max_abs_diff <- max(abs(heatmap_df$mean_difference_left_minus_right), na.rm = TRUE)
if (!is.finite(max_abs_diff) || max_abs_diff == 0) max_abs_diff <- 1

p_depth_heatmap <- ggplot(heatmap_df, aes(x = layer_depth, y = roi_name, fill = mean_difference_left_minus_right)) +
  geom_tile(color = "white", linewidth = 0.25) +
  geom_text(aes(label = ifelse(!is.na(p_fdr) & p_fdr < 0.05, "*", "")), size = 3) +
  facet_wrap(~ metric_label, nrow = 1) +
  scale_fill_gradient2(
    low = "#4575B4",
    mid = "white",
    high = "#D73027",
    midpoint = 0,
    limits = c(-max_abs_diff, max_abs_diff),
    name = "LH - RH"
  ) +
  labs(
    title = "Layer-Depth Hemisphere Differences Across Networks",
    x = "GPT-2 layer depth",
    y = "Network"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "right",
    strip.text = element_text(face = "bold")
  )

p_depth_heatmap

ggsave(
  filename = file.path(FIGURE_DIR, "layer_depth_lh_rh_mean_difference_heatmap.png"),
  plot = p_depth_heatmap,
  width = 14,
  height = 7,
  dpi = 300
)

p_effect_rank <- network_direction_summary %>%
  select(-analysis_note) %>%
  mutate(roi_name = fct_reorder(as.character(roi_name), mean_delta_lh_minus_rh_across_depths)) %>%
  ggplot(aes(x = mean_delta_lh_minus_rh_across_depths, y = roi_name, fill = network_family)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray55") +
  geom_col(width = 0.72, alpha = 0.9) +
  labs(
    title = "Network-Level LH/RH Difference In SRM-Related Gain",
    x = "Mean gain difference across layer depth (LH - RH)",
    y = "Network",
    fill = "Network family"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right")

p_effect_rank

ggsave(
  filename = file.path(FIGURE_DIR, "network_delta_lh_rh_effect_size_ranking.png"),
  plot = p_effect_rank,
  width = 11,
  height = 7,
  dpi = 300
)

if (exists("layer_model_results") && "F_value" %in% names(layer_model_results)) {
  model_plot_df <- layer_model_results %>%
    filter(value_col == "delta_mean_r") %>%
    mutate(
      roi_name = factor(roi_name, levels = NETWORK_ORDER),
      effect_label = factor(
        effect_label,
        levels = c("Hemisphere main effect", "Layer depth main effect", "Hemisphere x layer depth interaction")
      ),
      neg_log10_fdr = ifelse(is.na(p_fdr), NA_real_, -log10(pmax(p_fdr, 1e-300)))
    )

  p_model_heatmap <- ggplot(model_plot_df, aes(x = effect_label, y = roi_name, fill = neg_log10_fdr)) +
    geom_tile(color = "white", linewidth = 0.25) +
    geom_text(aes(label = significance), size = 3) +
    scale_fill_viridis_c(option = "magma", na.value = "gray90", name = "-log10(FDR p)") +
    labs(
      title = "Mixed-Effects Evidence For Hemisphere And Layer-Depth Effects",
      x = "Model effect",
      y = "Network"
    ) +
    theme_bw(base_size = 11) +
    theme(axis.text.x = element_text(angle = 35, hjust = 1), legend.position = "right")

  p_model_heatmap

  ggsave(
    filename = file.path(FIGURE_DIR, "network_lmer_effect_fdr_heatmap.png"),
    plot = p_model_heatmap,
    width = 11,
    height = 7,
    dpi = 300
  )
}
```

# 10. Final FDR-Significant Result Table

```{r final-significant-results}
model_sig <- if (exists("layer_model_results") && "F_value" %in% names(layer_model_results)) {
  layer_model_results %>%
    filter(!is.na(p_fdr), p_fdr < 0.05) %>%
    transmute(
      source_table = "network_wise_mixed_effects",
      analysis_question = model_question,
      result_type = "F-test",
      metric = value_col,
      metric_label,
      roi_name,
      network_family,
      layer_depth = NA_character_,
      layer_index = NA_integer_,
      source_effect = effect,
      effect = effect_label,
      direction = NA_character_,
      statistic_name = "F",
      statistic = F_value,
      df1 = NumDF,
      df2 = DenDF,
      p_value,
      p_fdr,
      significance,
      result_explanation = case_when(
        source_effect == "hemisphere:scale(log_layer_depth_percent)" ~
          "The LH/RH difference changes as GPT-2 layer depth changes.",
        source_effect == "hemisphere" ~
          "One hemisphere differs from the other after accounting for layer depth.",
        source_effect == "scale(log_layer_depth_percent)" ~
          "Encoding changes systematically with GPT-2 layer depth.",
        TRUE ~ "Significant mixed-effects model result."
      )
    )
} else {
  tibble()
}

welch_depth_sig <- layer_depth_welch_tests %>%
  filter(!is.na(p_fdr), p_fdr < 0.05) %>%
  transmute(
    source_table = "layer_depth_welch_tests",
    analysis_question = case_when(
      value_col == "srm_mean_r" ~ "Q1",
      value_col == "delta_mean_r" ~ "Q2",
      TRUE ~ NA_character_
    ),
    result_type = "Welch t-test",
    metric = value_col,
    metric_label,
    roi_name,
    network_family,
    layer_depth = as.character(layer_depth),
    layer_index = NA_integer_,
    effect = "LH vs RH",
    direction = case_when(
      mean_difference_left_minus_right > 0 ~ "LH > RH",
      mean_difference_left_minus_right < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    ),
    statistic_name = "t",
    statistic,
    df,
    df2 = NA_real_,
    p_value,
    p_fdr,
    significance,
    result_explanation = "Within this network and layer depth, LH and RH parcels differ in the tested encoding outcome."
  ) %>%
  rename(df1 = df)

perm_depth_sig <- layer_depth_permutation_tests %>%
  filter(!is.na(p_fdr), p_fdr < 0.05) %>%
  transmute(
    source_table = "layer_depth_permutation_tests",
    analysis_question = case_when(
      value_col == "srm_mean_r" ~ "Q1",
      value_col == "delta_mean_r" ~ "Q2",
      TRUE ~ NA_character_
    ),
    result_type = "Permutation test",
    metric = value_col,
    metric_label,
    roi_name,
    network_family,
    layer_depth = as.character(layer_depth),
    layer_index = NA_integer_,
    effect = "LH vs RH",
    direction = case_when(
      mean_difference_left_minus_right > 0 ~ "LH > RH",
      mean_difference_left_minus_right < 0 ~ "RH > LH",
      TRUE ~ "LH = RH"
    ),
    statistic_name = "mean difference",
    statistic = mean_difference_left_minus_right,
    df1 = NA_real_,
    df2 = NA_real_,
    p_value,
    p_fdr,
    significance,
    result_explanation = "Permutation confirms a layer-depth LH/RH difference without relying on Welch t-test assumptions."
  )

exact_welch_sig <- if (RUN_EXACT_LAYER_TESTS && "p_fdr" %in% names(exact_layer_welch_tests)) {
  exact_layer_welch_tests %>%
    filter(!is.na(p_fdr), p_fdr < 0.05) %>%
    transmute(
      source_table = "exact_layer_welch_tests",
      analysis_question = case_when(
        value_col == "srm_mean_r" ~ "Q1 localization",
        value_col == "delta_mean_r" ~ "Q2 localization",
        TRUE ~ NA_character_
      ),
      result_type = "Exact-layer Welch t-test",
      metric = value_col,
      metric_label,
      roi_name,
      network_family,
      layer_depth = as.character(layer_depth),
      layer_index,
      effect = "LH vs RH",
      direction = case_when(
        mean_difference_left_minus_right > 0 ~ "LH > RH",
        mean_difference_left_minus_right < 0 ~ "RH > LH",
        TRUE ~ "LH = RH"
      ),
      statistic_name = "t",
      statistic,
      df,
      df2 = NA_real_,
      p_value,
      p_fdr,
      significance,
      result_explanation = "Exploratory exact-layer localization; this identifies a specific GPT-2 layer where LH/RH parcels differ."
    ) %>%
    rename(df1 = df)
} else {
  tibble()
}

exact_perm_sig <- if (RUN_EXACT_LAYER_TESTS && "p_fdr" %in% names(exact_layer_permutation_tests)) {
  exact_layer_permutation_tests %>%
    filter(!is.na(p_fdr), p_fdr < 0.05) %>%
    transmute(
      source_table = "exact_layer_permutation_tests",
      analysis_question = case_when(
        value_col == "srm_mean_r" ~ "Q1 localization",
        value_col == "delta_mean_r" ~ "Q2 localization",
        TRUE ~ NA_character_
      ),
      result_type = "Exact-layer permutation test",
      metric = value_col,
      metric_label,
      roi_name,
      network_family,
      layer_depth = as.character(layer_depth),
      layer_index,
      effect = "LH vs RH",
      direction = case_when(
        mean_difference_left_minus_right > 0 ~ "LH > RH",
        mean_difference_left_minus_right < 0 ~ "RH > LH",
        TRUE ~ "LH = RH"
      ),
      statistic_name = "mean difference",
      statistic = mean_difference_left_minus_right,
      df1 = NA_real_,
      df2 = NA_real_,
      p_value,
      p_fdr,
      significance,
      result_explanation = "Exploratory exact-layer permutation localization; this is the non-parametric confirmation of an exact-layer LH/RH difference."
    )
} else {
  tibble()
}

final_fdr_significant_results <- bind_rows(
  model_sig,
  welch_depth_sig,
  perm_depth_sig,
  exact_welch_sig,
  exact_perm_sig
) %>%
  arrange(source_table, metric, p_fdr, roi_name, layer_index) %>%
  add_note(
    "This is the final summary table requested for reporting. It contains every result in this Rmd that survives FDR correction, along with a short explanation of what that result means."
  )

final_fdr_significant_results
```

# 11. Save Organized Workbook

```{r save-workbook}
table_notes <- tribble(
  ~sheet_name, ~analysis_question, ~what_this_sheet_does, ~why_it_matters,
  "00_README", "All questions",
  "Explains each sheet in the workbook and how the analyses map onto the three hemisphere questions.",
  "This is the first sheet so the workbook can be read without opening the Rmd.",
  "01_input_sanity", "Input validation",
  "Checks the number of parcel-layer rows, parcels, networks, layers, and layer-depth categories.",
  "This confirms that the statistical tests are based on the expected data structure.",
  "02_question_map", "All questions",
  "Defines Q1, Q2, and Q3 and links each question to its primary metric and statistical test.",
  "This keeps the analysis order aligned with the manuscript logic.",
  "03_outcome_definitions", "All questions",
  "Defines raw_mean_r, srm_mean_r, and delta_mean_r.",
  "This prevents ambiguity about what each encoding outcome means.",
  "04_hemisphere_counts", "Input validation",
  "Reports LH/RH parcel counts in each network.",
  "This helps interpret Welch and permutation tests when networks have unequal hemisphere parcel counts.",
  "05_profile_values", "Q1 and Q2",
  "Averages raw, SRM, and delta encoding values by network, hemisphere, and layer.",
  "This supports the layer-profile figures and provides descriptive values.",
  "06_mixed_models", "Q3",
  "Tests hemisphere, layer depth, and hemisphere-by-layer-depth effects within each network using mixed-effects models.",
  "This is the main model-based evidence for whether LH/RH effects change with GPT-2 layer depth.",
  "07_layer_depth_welch", "Q1 and Q2",
  "Compares LH and RH parcels within each network and layer-depth category using Welch tests.",
  "This is the first follow-up test for specific layer-depth categories.",
  "08_layer_depth_permutation", "Q1 and Q2",
  "Repeats the layer-depth LH/RH tests with label permutations.",
  "This provides a non-parametric robustness check after Welch tests.",
  "09_exact_layer_welch", "Q1/Q2 localization",
  "Exploratory Welch tests at each exact GPT-2 layer.",
  "This can localize which exact layers drive broader layer-depth effects.",
  "10_exact_layer_permutation", "Q1/Q2 localization",
  "Exploratory exact-layer permutation tests.",
  "This provides a non-parametric check for exact-layer localization.",
  "11_direction_summary", "Q1 and Q2",
  "Summarizes LH-RH direction across layer depth for each network.",
  "This helps describe whether a network is consistently LH- or RH-biased.",
  "12_final_FDR_results", "Final reporting",
  "Lists all FDR-significant results from mixed models, layer-depth tests, and exact-layer localization.",
  "This is the main report-ready summary table."
)

workbook_path <- file.path(TABLE_DIR, "final_r_test_organized_layerwise_hemisphere_results.xlsx")

writexl::write_xlsx(
  list(
    `00_README` = table_notes,
    `01_input_sanity` = input_sanity_summary,
    `02_question_map` = question_table,
    `03_outcome_definitions` = outcome_definition,
    `04_hemisphere_counts` = hemisphere_counts,
    `05_profile_values` = profile_df,
    `06_mixed_models` = layer_model_results,
    `07_layer_depth_welch` = layer_depth_welch_tests,
    `08_layer_depth_permutation` = layer_depth_permutation_tests,
    `09_exact_layer_welch` = exact_layer_welch_tests,
    `10_exact_layer_permutation` = exact_layer_permutation_tests,
    `11_layer_depth_direction` = layer_depth_direction_summary,
    `11_network_direction` = network_direction_summary,
    `12_final_FDR_results` = final_fdr_significant_results
  ),
  workbook_path
)

cat("\nFinal organized R analysis completed.\n")
cat("Workbook saved to:", workbook_path, "\n")
cat("Figures saved to:", FIGURE_DIR, "\n")
cat("Number of FDR-significant rows in final table:", nrow(final_fdr_significant_results), "\n")
```

# Interpretation Notes

- Positive `LH - RH` means the left hemisphere has larger values than the right hemisphere.
- `delta_mean_r` is the main SRM-related improvement measure: SRM+LLM encoding r minus Raw+LLM encoding r.
- `srm_mean_r` answers whether final SRM+LLM encoding is stronger in one hemisphere.
- `layer depth` refers to the GPT-2 representational depth category: embedding, early, early-middle, middle-late, or late.
- Welch tests are always reported before permutation tests.
- Mixed-effects models are the main analysis for whether hemisphere effects change continuously with GPT-2 layer depth.
